---
title: "Assignment 3_ML"
author: "Lukas van der Watt"
date: "10/16/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Calling packages required to run the various commands
library(e1071)
library(caret)
UB.data <- read.csv("UniversalBank.csv") #reading the .csv file for universal bank data
str(UB.data) #Looking at the structure of the data to see what kind of variables are present.
```
```{r}
#converting to factors
UB.data$Personal.Loan <- as.factor(UB.data$Personal.Loan) #Converting the Personal Loan variable to as.factor
UB.data$Online <- as.factor(UB.data$Online)#Converting the Online variable to as.factor
UB.data$CreditCard <- as.factor(UB.data$CreditCard)#Converting the CreditCard variable to as.factor
str(UB.data)
```
```{r}
#Partition the data into training(60) validation(40)
selected.var <- c(10,13,14) #Selecting variables to be partitioned

set.seed(123) #randomize
train.in <- createDataPartition(UB.data$Online,p = 0.6, list = FALSE) #creating a training index with the stratification variable (online) that contains 60% of the Universal bank data
ub.data.train <- UB.data[train.in,selected.var] #Training set
ub.data.valid <- UB.data[-train.in,selected.var] #Validation set
str(ub.data.train) #structure of the training set
```
PART A : Creating a Pivot Table
```{r}
#Method 1: This shows the pivot table with row variables (Online and CreditCard) and the column variables(Personal.Loan) This table shows the count
attach(ub.data.train) # Attaching the training set to the following statements
ftable(Personal.Loan, CreditCard, Online) #Creating a pivot table with personal loan as a column variable and the Online and CreditCard
detach(ub.data.train)
```
Part B Ans. P(Personal.Loan=1 | CreditCard=1 , Online=1) = 47/(47+478) = 0.08952381
```{r}
#Method 2: This shows the pivot table with row variables (Personal Loan and CreditCard) and the column variables(Online). This prop table shows the probabilities 
prop.table(ftable(ub.data.train, row.vars = c(1,3), column.vars = 2 ),margin = 1)
```

PART C
```{r}
#Creating two seperate pivot tables 
attach(ub.data.train)
prop.table(table(Personal.Loan, Online), margin = 1)
prop.table(table(Personal.Loan, CreditCard), margin = 1)
detach(ub.data.train)

```

i.   P(CC = 1 | Loan = 1) (CC=1 means CreditCard holder, Loan=1 means Loan was accepted)
          Ans. = 0.30249
          
ii.  P(Online = 1 | Loan = 1)
          Ans. = 0.59075
          
iii. P(Loan = 1) (the proportion of loan acceptors)
          Ans. = (77+119+38+47)/3001 = 0.09363545
          
iv.  P(CC = 1 | Loan = 0)
          Ans. = 0.2893382
          
v.   P(Online = 1 | Loan = 0)
          Ans. = 0.5974265
          
vi.  P(Loan = 0)
          Ans. = (786+309+1147+478)/3001 = 0.9063645
          or   = 1-P(Loan=1) = 1-0.09363545 =  0.9063646

PART E:  P(Loan = 1 | CC= 1, Online = 1) 
= [P(CC=1 , Online=1 | Loan = 1)*P(Loan=1)]/P(CC=1 , Online=1)

= [P(CC=1 | Loan=1) * P(Online=1 | Loan=1)* P(Loan=1)]/[P(CC=1 | Loan=1) * P(Online=1 | Loan=1) * P(Loan=1)
+P(CC=1 | Loan=0) * P(Online=1 | Loan=0)*P(Loan=0)]

=((0.3024911) * (0.59075) * (0.09363545))/(((0.3024911) * (0.59075) * (0.09363545))+((0.2893382) * (0.5974265) * (0.9063645)))

= 0.09649284


*******PART F: The Results between the Pivot table and calculated Naive Bayes value differs. For the Naive Bayes a value probability of 0.09649284 was achieved whereas the Pivot Table indicated a probability of 0.08952381 for the P(Loan=1 | CC=1 , Online=1) statement. The Naive Bayes calculation is higher than the value in the Pivot table. The Naive Bayes calculation is not exact as we are making assumptions. Only the numerator is approximated in this calculation. The probability value in the pivot table is therefore more accurate.

PART G:
```{r}
UBdata.nb <- naiveBayes(Personal.Loan~., data = ub.data.train)# Creating the NAivive Bayes Model on the training set
UBdata.nb
```
Naive Bayes Calculation of P(Loan = 1 | CC = 1, Online = 1) = [P(CC=1 , Online=1 | Loan = 1)*P(Loan=1)]/P(CC=1 , Online=1)

= ((0.3024911) * (0.5907473) * (0.09363545)) / (((0.3024911) * (0.5907473) * (0.09363545))+((0.2893382) * (0.5974265) * (0.90636455)))

= 0.09649244

Number in E = 0.09649284

As we can see the values are essentially the same between Part E and Part G values.

